{\bf Systematic Testing: A New Perspective on Races}

### 2014-12-02 landslide-lecture/2014/lecture.tex
add in progress 410 lecture for f14

\begin{frame}{Testing}
	How can programmers be confident in their code's correctness?
	\begin{itemize}
		\item Theorem proving
		\item Unit tests
		\item Stress tests
		\begin{itemize}
			\item (state of the art in 15-410)
		\end{itemize}
		\item Releasing to paying customers and worrying about correctness later
	\end{itemize}
	\linegap

@@

% TODO TODO 2014: Move this slide to after "systematic testing", before "more on preemption points"
\begin{frame}{Preemption Points} % TODO: duplicate slide with later
	A {\bf preemption point} is\ldots

	\linegap
	A code location where being preempted causes different behaviour.
	\begin{itemize}
		\item Intuitively: Somewhere that interesting interleavings can happen around.
	\end{itemize}
	Examples:
	\begin{itemize}
		\item A new thread becomes runnable
		\item Voluntary reschedule (e.g. \texttt{yield}, \texttt{cond\_wait})
		\item Synchronization primitives
	\end{itemize}
	% Say: We'll revisit this later.
\end{frame}

@@

\begin{frame}{Systematic Testing}
	\textbf{Systematic testing} is:
	\begin{itemize}
		\item Systematically enumerating different interleavings
		\begin{itemize}
			\item Intuitively: Generate many ``tabular execution traces''
		\end{itemize}
		\item Exploring all branches in these trees
		\begin{itemize}
			\item (by controlling scheduling decisions at preemption points)
		\end{itemize}
		\item In practice: Depth-first search of branches
	\end{itemize}
\end{frame}

% TODO TODO 2014: reorg subsection dividers
\subsection{Requirements}

\begin{frame}{More on Preemption Points}
	Important point: What does ``all possible interleavings'' mean?

	\linegap
	One extreme: Decide at every instruction
	\begin{itemize}
		\item {\em Good news}: Will find every possible race condition.
		\item {\em Bad news}: Runtime of test will be impossibly large.
	\end{itemize}
	\linegap

@@

\begin{frame}{More on Preemption Points}
	Sweet spot: Insert a thread switch everywhere it ``might matter''.

	\linegap
	When do we fear being preempted?
	\begin{itemize}
		\item New threads becoming runnable (\texttt{fork}, \texttt{cond\_signal}, etc)
			\begin{itemize}
				\item Preemptions may cause it to run before we're ready
			\end{itemize}
		\item Synchronization primitives (\texttt{mutex\_lock}, etc)
			\begin{itemize}
				\item If used improperly\ldots
			\end{itemize}
		\item Unprotected shared memory accesses % Say: more on this later.
			\begin{itemize}
				\item May result in data structure corruption
			\end{itemize}
	\end{itemize}
	\linegap

	Finding the sweet spot is a joint effort between programmer and tool. (More on this later.)
\end{frame}

@@

>>>>

\begin{frame}{Testing}
	How can programmers be confident in their code's correctness?
	\begin{itemize}
		\item Unit tests
		\begin{itemize}
			\item good for basic functionality, bad for concurrency
		\end{itemize}
		\item Stress tests
		\begin{itemize}
			\item state of the art in 15-410
		\end{itemize}
		\item Theorem proving
		\begin{itemize}
			\item heavy burden on the programmers
		\end{itemize}
		\item Releasing to paying customers and worrying about correctness later
	\end{itemize}
	\linegap

@@

@@

\begin{frame}{Systematic Testing - The Big Picture}
	\textbf{Goal: Force the system to execute every possible interleaving.}
	% Say: checking for crashes/memory errors/etc in each one.
	\begin{itemize}
		\item At end of each branch, rewind system and restart test.
		\item Artificially add preemptions to test different thread interleavings.
		\item Intuitively: Generate many ``tabular execution traces''.
	\end{itemize}
	\pause
	\linegap

	Okay, wait a sec...
	\begin{itemize}
		\item How can you possibly execute {\em every possible} interleaving?
		\item How can you know where preempting matters most?
	\end{itemize}
\end{frame}

\begin{frame}{Preemption Points}
	\textbf{Preemption points} (PPs) are code location where being preempted may cause different behaviour.
	\begin{itemize}
		\item IOW, somewhere that interesting interleavings can happen around.
	\end{itemize}
	\linegap

	Systematic tests are {\em parameterized} by the set of PPs.
	\begin{itemize}
		\item If there are $n$ PPs and $k$ threads, state space size is $n^k$.
		\item Need to choose the set of PPs very carefully for test to be effective.
	\end{itemize}
\end{frame}

\begin{frame}{Preemption Points}
	What does ``all possible interleavings'' actually mean?

	\linegap
	One extreme: Preempt at every instruction
	\begin{itemize}
		\item {\em Good news}: Will find every possible race condition.
		\item {\em Bad news}: Runtime of test will be impossibly large.
	\end{itemize}
	\linegap

@@

\begin{frame}{Preemption Points}
	Sweet spot: Insert a thread switch everywhere it ``might matter''.

	\linegap
	When do we fear being preempted?
	\begin{itemize}
		\item Threads becoming runnable (\texttt{fork()}, \texttt{cond\_signal()}, etc.)
			\begin{itemize}
				\item Preemptions may cause it to run before we're ready
			\end{itemize}
		\item Synchronization primitives (\texttt{mutex\_lock()}/\texttt{unlock()}, etc.)
			\begin{itemize}
				\item If used improperly\ldots
			\end{itemize}
		\item Unprotected shared memory accesses % Say: more on this later.
			\begin{itemize}
				\item May result in data structure corruption
			\end{itemize}
	\end{itemize}
	\linegap

	%%%% Not anymore! :D
	%Finding the sweet spot is a joint effort between programmer and tool. (More on this later.)
\end{frame}

@@

%% TODO -- left off working on slides here.
